---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# permuco4brain

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/jaromilfrossard/permuco4brain.svg?branch=master)](https://travis-ci.org/jaromilfrossard/permuco4brain)
<!-- badges: end -->

`permuco4brain` provides functions to compute permutation test in brain imagery data. It is specially designed for M-EEG/ERP data. `permuco4brain` is an add-on to `permuco` to computing clustermass tests on spatio-temporal data (eg: full-scalp EEG data).

## Installation

`permuco4brain` currently works with `permuco` 1.1.1 (github). Download it using: 

```{r eval=FALSE, warning=FALSE, include=TRUE}
devtools::install_github("jaromilfrossard/permuco", build_vignettes = TRUE)
```

`permuco4brain` is still under development and you can expect changes. Make sure to install `permuco4brain` with its full documentation:

```{r eval=FALSE, warning=FALSE, include=TRUE}
devtools::install_github("jaromilfrossard/permuco4brain", build_vignettes = TRUE)
```

## Functionality

The main function of `permuco4brain` is `brainperm()` and it needs:

1. `formula`: a formula object defining the design (right side) and a 3D array containing the signals (left side). The signal is a 3D array visible in the global environment. Its dimension is the design X samples X channels. Each "rows" of the 3D array should be related to the corresponding row of the design (`data` argument).
2. `data`: a data-frame containing the variables of the design.
3. `graph`: an `igraph` object defining spatial adjacency of the channels (third dimension of the signal). The names of the vertices should correspond to the names of the 3rd dimension of signal.

You can inspect result using the `summary()` method and visualize them using `image()`.


## Documentation

Visite https://jaromilfrossard.github.io/permuco4brain or check the vignette:

1. The [permuco4brain-with-eeguana](https://jaromilfrossard.github.io/permuco4brain/articles/permuco4brain-with-eeguana.html) vignette presents how to use `permuco4brain` in combination with `eeguana` :

```{r eval=FALSE, warning=FALSE, include=TRUE}
vignette("permuco4brain-with-eeguana", package = "permuco4brain")
```

You will find information how to extract the 3D array, the design data-frame and the graph from an `eeg_lst` object of the `eeguana` package. Check the `eeguana` package for pre-processing EEG data within `R` (https://github.com/bnicenboim/eeguana).


2. The [download-example-cheval](https://jaromilfrossard.github.io/permuco4brain/articles/download-example-cheval.html) vignette presents how to use `permuco4brain` with signals stored in `edf` files:

```{r eval=FALSE, warning=FALSE, include=TRUE}
vignette("download-example-cheval", package = "permuco4brain")
```

In this tutorial, you learn how to download EEG data from `zenodo` and analyze them using `permuco4brain`.


3. The [figure-ggplot2](https://jaromilfrossard.github.io/permuco4brain/articles/figure-ggplot2.html) vignette presents how to use `ggplot2` to produce graphical representation for publication:

```{r eval=FALSE, warning=FALSE, include=TRUE}
vignette("figure-ggplot2", package = "permuco4brain")
```

In this tutorial, you find 3 different types of plot that present the results of a clustermass test. All of them use the `ggplot2` package and may easily be customized for publication.

