---
title: "Download example of full scalp EEG data"
author: "Jaromil Frossard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ref.bib
vignette: >
  %\VignetteIndexEntry{download-example-cheval}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
library(rgl)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knit_hooks$set(webgl = hook_webgl)
```



In the following tutorial, we download data analysed by [@cheval]. They are stored in the [zenodo repository](https://zenodo.org/record/1169140) in a "edf" files.

## Installing packages

We need the several packages including `tidyverse`  [@tidyverse], `edf` [@edf], `abind` [@abind], `xlsx` [@xlsx] and `permuco4brain` [@frossard_permuco4brain] that we active using:

```{r package, echo=TRUE,eval=FALSE}
library(tidyverse)
library(edf)
library(abind)
library(xlsx)
library(permuco4brain)
```

## Downloading data

We download the data in a temporary files and we unzip the edf files also in a temporary file.

```{r download, echo=TRUE,eval=FALSE}
temp_zip <- tempfile()
temp_edf <- tempfile()
download.file("https://zenodo.org/record/1169140/files/ERP_by_subject_by_condition_data.zip",temp_zip)
unzip(zipfile = temp_zip, exdir = temp_edf)
```


We store the names of the files in a data frame using:

```{r filename, echo=TRUE,eval=FALSE}
df <- data.frame(filename = list.files(paste0(temp_edf,"/raw_data/")),
                 stringsAsFactors = F)
```


If your are not familiar with edf files you can explore the first file using:

```{r edf, echo=TRUE,eval=FALSE}
edf1 <- read.edf(file.path(temp_edf,paste0("raw_data/",list.files(paste0(temp_edf,"/raw_data/"))[1])))
```


We download information about the subject and we keep only the meaure "Moderate to Vigorous Physical Activity" (MVPA) that we center:

```{r beetween, echo=TRUE,eval=FALSE}
df_bw <- read.csv("https://zenodo.org/record/1169140/files/data_self_report_R_subset_zen.csv",sep=";")
df_bw <- df_bw %>%
  mutate(subject = paste0("S",subject),
         mvpa_c = MVPA-mean(MVPA))%>%
  select(subject,mvpa_c)
``` 


## Reshaping data

We create the design from the filenames and we join the "between-participant" information. Moreover we keep only neutral shapes "Round" rows.

```{r join, echo=TRUE,eval=FALSE}
df<- 
  df%>%
  group_by(filename)%>%
  separate(filename,into = c("participant","neutral_stim","av","action","shape","edf"),remove = F)%>%
  select(-edf,-av)%>%
  left_join(df_bw,by=c("participant"="subject"))%>%
  filter(!shape%in%c("Carre","Neutre"))
```


We import the edf files and retrive only the signals:

```{r signal, echo=TRUE,eval=FALSE}
df<-
  df%>%
  mutate(edf = map (filename,function(ii){
    path <- file.path(temp_edf,paste0("raw_data/",ii))
    read.edf(path)
  }))%>%
  mutate(signal = map(edf,function(listi){
    listi$signal%>%
      map(function(channel)channel$data)%>%
      as.data.frame()%>%
      as.matrix()
  }))%>%
  ungroup()
```

We compute the signal relative to the neutral "Round" shapes:

```{r relative, echo=TRUE,eval=FALSE}
df<-
  df%>%
  select(participant,shape,action,mvpa_c,signal)%>%
  pivot_wider(names_from = shape,values_from = signal)%>%
  mutate(PA_rel = map2(AP,Rond,`-`),
         Sed_rel = map2(SED,Rond,`-`))%>%
  select(-AP,-Rond,-SED)%>%
  pivot_longer(PA_rel:Sed_rel,names_to = "stimuli", values_to = "signal")
```

We `unlink()` the temporary files:

```{r unlink, echo=TRUE,eval=FALSE}
unlink(temp_zip)
unlink(temp_edf)
```

## Extracting the signal, the design and the graph

We create the 3D array that stores the signals:

```{r array, echo=TRUE,eval=FALSE}
signal<- 
  (df%>%
  pull(signal)%>%
  invoke(abind::abind,.,along = 3)%>%
  aperm(c(3,1,2)))[,102:512,]
```

We create the dataframe that stores the design:

```{r design, echo=TRUE,eval=FALSE}
design<-
  df%>%
  select(participant, action, stimuli,mvpa_c)%>%
  as.data.frame()
```

Download the position of the channel from [biosemi](www.biosemi.com):

```{r biosemi, echo=TRUE,eval=FALSE}
temp_xl <- tempfile()
download.file("https://www.biosemi.com/download/Cap_coords_all.xls",temp_xl,mode="wb")
coord <-  read.xlsx(file=temp_xl, sheetIndex = 3, header =T,startRow=34)
unlink(temp_xl)
```

We clean the data from biosemi and sort it accordingly to the thrid dimension of the signal array:

```{r coord, echo=TRUE,eval=FALSE}
coord<- 
  coord%>%
  select(channel = Electrode, x= "x...r.sinÎ..cosÏ.",
         y = "y...r.sinÎ..sinÏ.", z = "z...r.cosÎ.")%>%
  filter(!is.na(x),!is.na(y),!is.na(z))%>%
  mutate(channel = as.character(channel))%>%
  mutate(channel = if_else(channel=="Iz (inion)","Iz",channel),
         channel = if_else(channel=="T7 (T3)","T7",channel),
         channel = if_else(channel=="T8 (T4)","T8",channel),
         channel = if_else(channel=="Afz","AFz",channel))%>%
  arrange(match(channel,dimnames(signal)[3]))
```

The graph is computed using $\delta=35$ as the maximal distance defining adjacency:

```{r graph, echo=TRUE,eval=FALSE}
graph <- position_to_graph(coord,delta = 35, name="channel")
```


## Final step

The 3 objects `signal`, `design`, `graph` are saved for future re-use:

```{r save, echo=TRUE,eval=FALSE}
# save(signal,design,graph,file = "signal_design_graph.RData")
```

Or we simply run the permutation test:

```{r brainperm, echo=TRUE,eval=FALSE}
clustermass <- brainperm(signal ~ action*stimuli*mvpa_c + Error(participant/(action*stimuli)),
                         data = design, graph = graph)
```

# Reference

