---
title: "Using `permuco4brain` with `eeguana`"
author: "Jaromil Frossard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ref.bib
vignette: >
  %\VignetteIndexEntry{Using `permuco4brain` with `eeguana`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Extracting 3D array using eeguana

The `eeguana` package [@nicenboim_eeguana_2020] provides tools for the pre-processing of EEG data. Once the pre-processing is done with `eeguana`, the 3D matrix/array of the signals should be extracted from the `eeguana`. This signals are stored in a 3D matrix/array that is the left part of the formula of the `brainperm()` function to perform the cluster-mass test or TFCE.

The EEG data [@nicenboim_eeg_2019] available on the Open Science Framework repository was analyzed using the `eeguana` package by @nicenboim2019words. We use the same data-set in the following tutorial.

### Downloading data from OSF

From the [EEG data](https://osf.io/ut7xq/) repository we found the `"03_data_summarization.Rmd"` file which provides functions to easily download the EEG data. As it is a large dataset (more than 20GB), we only download the pre-processed signals using the [ICA](https://en.wikipedia.org/wiki/Independent_component_analysis). When following the first chunk of the `"03_data_summarization.Rmd"` file you should have (at least):

1. A folder `data/` containing 5 csv files, 2 tsv files and 2 folders. 
2. The folder `data/preproc_ica/` contains 120 RDA files (1 for each subject). Each file contains an `eeg_lst` object from `eeguana` with pre-processed EEG data.
3. The folder `data/opensesame_files/` contains 120 csv files (1 for each subject)


## R package

In the following analysis, we mainly use functions from the `tidyverse` [@tidyverse], from `eeguana` [@nicenboim_eeguana_2020], from `permuco` [@frossard_permuco], from `permuco4brain` [@frossard_permuco4brain], from `future` [@future] and from `igraph` [@csardi_igraph]:

```{r package, echo=TRUE,eval=FALSE}
library(tidyverse)
library(eeguana)
library(permuco)
library(permuco4brain)
library(igraph)
library(future)
```


## Computing the ERPs

First, we compute ERPs following the design of the experiment and save them in the folder `data_erp/`. To reduce the number of samples (and the computing time), we delete samples (or time-point) before the event and downsample to 256Hz. Finally, the ERPs are computed by taking the average for each sample, each channel and each experimental condition, which are the interaction between the `within` participant factors `region` and `cond`.



```{r within, echo=TRUE,eval=FALSE}
datadir <- "data/preproc_ica/"

lf <- list.files(datadir)

for(i in 1:length(lf)){
  print(i)
  eeg <- readRDS(paste0(datadir,lf[i]))
  
  erp <- eeg%>%
    filter(.sample>=0)%>%
    eeg_downsample(.q=2)%>%
    group_by(.recording,.sample,region,cond)%>%
    summarize_at(channel_names(.), mean, na.rm = TRUE)
  saveRDS(erp, file = paste0("data_erp/",lf[i]))
}
```

The previous script should produce 120 files in the folder `data_erp/`, one for each participant.

The ERP can be combine into 1 `eeg_lst` object by running:

```{r erps, echo=TRUE,eval=FALSE}
datadir <- "data_erp/"
lf <- list.files(datadir)

erps <- list()

for(i in 1:length(lf)){
  print(i)
  erps[[i]] <- readRDS(paste0(datadir,lf[i]))%>%
    mutate(subject_nr = as.numeric(strsplit(.recording,"[._]")[[1]][1]))
}

erps <- do.call("bind",erps)
```

Now, the `erps` object contains the ERPs for all participants. It is  `eeg_lst` object which can be manipulated using the functions from `eeguana`.

## Adding the `between`-participant variables

We download the `between`-participant variables by combining 4 files. We choose to keep the age, the batch and the occupation as `between`-participant factors. 

```{r between, echo=TRUE,eval=FALSE}
df_subj <- bind_rows(read_csv("data/students.csv") %>%
                         mutate(batch =1, student = "yes"),
                       read_csv("data/students_2.csv") %>%
                         mutate(batch =2, student = "yes"),
                       read_csv("data/nonstudents.csv") %>%
                         mutate(batch =1, student = "no"),
                       read_csv("data/nonstudents_2.csv")%>%
                         mutate(batch =2, student = "no"))%>%
  select(subj_order = NR, subject_nr, age = Alter, batch, 
         occupation = student)
```


The `.csv` data does not contain the `between`-participants variables for all participants, so we choose to delete 8 participants from our analysis. 


```{r remove, echo=TRUE,eval=FALSE}
subject_to_rm <- c(21,35,316,111,222,322,98,245)

erps <- 
  erps%>%
  filter(!(subject_nr%in%subject_to_rm))

df_subj <- 
  df_subj%>%
  filter(!(subject_nr%in%subject_to_rm))
```

Finally, we merge the `between`-participants variables to the `erps` object and also remove the signals from the ocular and reference channels the `erps` object.

```{r merge, echo=TRUE,eval=FALSE}
erps <- 
  erps%>%
  left_join(df_subj ,by = c("subject_nr"))

chan_to_rm <- c("M1","M2","HEOG","VEOG")
erps <- 
  erps%>%
  select(-one_of(chan_to_rm))
```

## Extracting the 3D matrix

We extract the 3D matrix using the following script. The dimensions are the design $\times$ the samples $\times$ the channels.

```{r array, echo=TRUE,eval=FALSE}
signal <- 
  erps%>%
  signal_tbl()%>%
  group_by(.id)%>%
  nest()%>%
  mutate(data = map(data,~as.matrix(.x[-1])))%>%
  pull(data)%>%
  invoke(abind::abind,.,along = 3)%>%
  aperm(c(3,1,2))
```

The design which contains both the between and within variables is simply extracted using the `segments_tbl()` function. Moreover, we center the covariate as it will make more sense for the test.

```{r design, echo=TRUE,eval=FALSE}
design <- 
  segments_tbl(erps)%>%
  ungroup()%>%
  mutate(age_c = age - mean(age))
```


Finally the graph of adjacency is computed using the 32 channels layout data in `eeguana` and the `position_to_graph()` function from `permuco4brain`. Moreover, we can check the adjacency graph using the `plot()` function from `igraph`.

```{r graph, echo=TRUE,eval=FALSE}
layout <- 
  channels_tbl(erps)

graph <- position_to_graph(layout, delta = .75,name = .channel,
                          x = .x, y = .y, z = .z)
```

You can use `igraph::rglplot(graph)` to have a 3D representation of the adjacency between channels.
Before running the clustermass test, you may want to save the 3 objects: the 3D array containing the signals, the dataframe containing the design and the graph defining the spatial adjacency of the channels.

```{r save, echo=TRUE,eval=FALSE}
# save(signal = signal,
#      design = design,
#      graph = graph,file = "signal_design_graph.RData")
```

Finally, we use `permuco4brain` for testing the difference between experimental conditions. We have a repeated measures ANCOVA, with 1 covariate (the centered age), 1 between-participants factors (the occupation) and 2 within-participants factors (region and condition). The permutation tests is computed for all samples, all channels while controlling for the family-wise error rate (FWER).

The `plan()` function from the [`future` package](https://github.com/HenrikBengtsson/future) handles multi-cores computing


```{r model, echo=TRUE,eval=FALSE}
formula <- signal ~ age_c*occupation*region*cond +Error(.recording/(region*cond))
plan(multisession)
model <- brainperm(formula = formula, data = design, graph = graph)
```

By default, the `brainperm()` function run the cluster-mass test [@maris_nonparametric_2007]. However, the procedure from @troendle_stepwise_1995 and the TFCE [@smith_thresholdfree_2009] are also available by specifying the arguments `multcomp = "troendle"` or `multcomp = "tfce"`.

### Reference

