---
title: "Using `permuco4brain` with `eeguana`"
author: "Jaromil Frossard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ref.bib
vignette: >
  %\VignetteIndexEntry{Using `permuco4brain` with `eeguana`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Extracting 3D array using eeguana

The `eeguana` package [@nicenboim_eeguana_2020] provides tools for the pre-processing of EEG data. Once the pre-processing done with eeguana the 3D matrix can be easily extracted to be pass in the `brainperm()` function.

The EEG data [@nicenboim_eeg_2019] available on an Open Science Framework repository was analysed using the `eeguana` package by @nicenboim2019words. We use the same data-set for the following tutorial.

### Downloading data from OSF

From the [EEG data](https://osf.io/ut7xq/) repository we found the "03_data_summarization.Rmd" file which provides functions to easily download the data-set. It is large (more than 20Gb) so we only download the pre-processed data using ICA. Following the first chunk of the "03_data_summarization.Rmd" file you should have (at least):

1. A folder `data/` containing 5 csv files, 2 tsv files and 2 folders. 
2. The folder `data/preproc_ica/` contains 120 RDA files (1 for each subject). Each file contains an `eeg_lst` object from eeguana with pre-processed EEG data.
3. The folder `data/opensesame_files/` contains 120 csv files (1 for each subject)


## R package

In the following analysis, we will mainly use functions from the `tidyverse` [@tidyverse], from `eeguana` [@nicenboim_eeguana_2020], from `permuco` [@frossard_permuco], from `permuco4brain` [@frossard_permuco4brain] and from `igraph` [@csardi_igraph]:

```{r package, echo=TRUE,eval=FALSE}
library(tidyverse)
library(eeguana)
library(permuco)
library(permuco4brain)
library(igraph)
```


## Computing the ERPs

First, we compute ERP following the design of the experiment and save them in the folder `data_erp/`. To reduce the number of sample (and the computing time), we delete samples (or time-point) before the event and down-sample to 256Hz. Finally the ERPs are computed by taking the average for each sample, each channel and each experimental condition, which are the interaction between the `within` participant factors `region` and `cond`.





```{r within, echo=TRUE,eval=FALSE}
datadir <- "data/preproc_ica/"

lf <- list.files(datadir)

for(i in 1:length(lf)){
  print(i)
  eeg <- readRDS(paste0(datadir,lf[i]))
  
  erp <- eeg%>%
    filter(.sample>=0)%>%
    eeg_downsample(q=2)%>%
    group_by(.recording,.sample,region,cond)%>%
    summarize_at(channel_names(.), mean, na.rm = TRUE)
  saveRDS(erp, file = paste0("data_erp/",lf[i]))
}
```

The previous script should produce 120 files in the folder `data_erp/`, one for each participant.

The ERP can be combine into 1 `eeg_lst` object by running:

```{r erps, echo=TRUE,eval=FALSE}
datadir <- "data_erp/"
lf <- list.files(datadir)

erps <- list()

for(i in 1:length(lf)){
  print(i)
  erps[[i]] <- readRDS(paste0(datadir,lf[i]))%>%
    mutate(subject_nr = as.numeric(strsplit(.recording,"[._]")[[1]][1]))
}

erps <- do.call("bind",erps)
```

Now, the `erps` object contains the ERP for all participants. Its class is `eeg_lst` which means that it can be manipulate using the functions of `eeguana`.

## Adding the `between`-participant variables

We download the `between`-participant variables by combining 4 files. We choose to keep the age, the batch and the occupation as `between`-participant factors. 

```{r between, echo=TRUE,eval=FALSE}
df_subj <- bind_rows(read_csv("data/students.csv") %>%
                         mutate(batch =1, student = "yes"),
                       read_csv("data/students_2.csv") %>%
                         mutate(batch =2, student = "yes"),
                       read_csv("data/nonstudents.csv") %>%
                         mutate(batch =1, student = "no"),
                       read_csv("data/nonstudents_2.csv")%>%
                         mutate(batch =2, student = "no"))%>%
  select(subj_order = NR, subject_nr, age = Alter, batch, 
         occupation = student)
```


The data-set does not contain the `between`-participants variables for all participants, so we choose to delete 8 participants from our analysis. 


```{r remove, echo=TRUE,eval=FALSE}
subject_to_rm <- c(21,35,316,111,222,322,98,245)

erps <- 
  erps%>%
  filter(!(subject_nr%in%subject_to_rm))

df_subj <- 
  df_subj%>%
  filter(!(subject_nr%in%subject_to_rm))
```

Finally we merge the `between`-participants variables to the `erps` object and remove the uninteresting channels from our analysis.

```{r merge, echo=TRUE,eval=FALSE}
erps <- 
  erps%>%
  left_join(df_subj ,by = c("subject_nr"))

chan_to_rm <- c("M1","M2","HEOG","VEOG")
erps <- 
  erps%>%
  select(-one_of(chan_to_rm))
```

## Extracting the 3D matrix

We extract the 3D matrix using the following script. The dimensions are design $\times$ sample $\times$ channel.

```{r array, echo=TRUE,eval=FALSE}
signal <- 
  erps%>%
  signal_tbl()%>%
  group_by(.id)%>%
  nest()%>%
  mutate(data = map(data,~as.matrix(.x[-1])))%>%
  pull(data)%>%
  invoke(abind::abind,.,along = 3)%>%
  aperm(c(3,1,2))
```

The design which contains both the between and within variables is simply extracted using the `segments_tbl()` function. Moreover, we center the covariate as it makes more sense for the test.

```{r design, echo=TRUE,eval=FALSE}
design <- 
  segments_tbl(erps)%>%
  ungroup()%>%
  mutate(age_c = age - mean(age))
```


Finally the graph of adjacency is computed using the 32 channels layout data in `eeguana` and the `position_to_graph()` function from `permuco4brain`. Note that, as previously we deleted some channels. Moreover, we can check the adjacency graph using the `plot()` function from `igraph`.

```{r graph, echo=TRUE,eval=FALSE}
layout <- 
  channels_tbl(erps)%>%
  filter(!(.channel%in%c(chan_to_rm)))

graph <- position_to_graph(layout, delta = .75,name = .channel,
                          x = .x, y = .y, z = .z)
```

You can use `igraph::rglplot(graph)` to have a 3D representation of the adjacency between channels.
Now, we have the 3 object we need to run a clustermass test, the 3D array containing the signals, a data-frame containing the design and a graph defining the spatial adjacency of the channels. We can save these object before running the analysis.

```{r save, echo=TRUE,eval=FALSE}
# save(signal = signal,
#      design = design,
#      graph = graph,file = "signal_design_graph.RData")
```

Finally, we use `permuco4brain` for testing the difference between experimental conditions. We have a repeated measures ANCOVA, with 1 covariate (the centered age), 1 between-participants factors (the occupation) and 2 within-participants factors (reguion and conditions). The permutation tests is computed for all samples, all channels while controlling for the family-wise error rate (FWER).


```{r model, echo=TRUE,eval=FALSE}
formula <- signal ~ age_c*occupation*region*cond +Error(.recording/(region*cond))
model <- brainperm(formula = formula, data = design, graph = graph)
```


### Bibliography

